### 2021-04-15 manuscript.Rmd
Typos and tifs.

* __Specimen Collection, Processing, and RT-PCR Assay__: Specimens were collected during routine clinical practice using a nylon flocked mini-tip swab collected in VTM or saline [@Garnett2020-dl; @Rogers2020-di; @Richard-Greenblatt2021-ql; @Rodino2020-dk]. Healthcare providers obtained samples using CDC guidelines for NP samples collection. Samples were transported to the laboratory at ambient temperature and stored at 4⁰C if not run immediately. All specimens were analyzed at the Hospital of the University of Pennsylvania's Clinical Microbiology Laboratory using an EUA-approved laboratory-developed SARS-CoV-2 assay on the BD MAX system. The multiplex assay was designed to include a specimen quality control (β-actin)[@Singanayagam2020-mg; @Smieja2010-op], an internal processing control to monitor the RT and PCR steps (MS2 phage RNA)[@Dreier2005-ye], and a SARS-CoV-2 target (Orf1ab). Exk TNA2 extraction reagent kits (Becton Dickinson) for the BD MAX open system reagent suite were used for the lab-developed SARS-CoV-2 assay based on a previously described assay [@LeBlanc2020-wa; @Vermeiren2020-pm]. The BD MAX system was set to run type 1 workflow. PCR conditions consisted of a reverse transcriptase step (600s at 58⁰C, 1 cycle), denaturation step (60s at 98⁰C, 1 cycle) and extension steps (10s at 98⁰C followed by 40s at 63⁰C, 40 cycles). Ct values for all three targets (β-actin, MS2 phage RNA) were recorded. <span style='background-color:lightcyan'>The acceptable range for β-actin in our assay was 24-37 Cts. Samples that were negative for SARS-CoV-2 and either negative for β-actin or with a β-actin Ct > 37 were reported as invalid.</span>

@@

* __SARS-CoV-2 Detection and Cycle Threshold Range__: Of 1282 tested specimens, 134 were found to have detectable SARS-CoV-2 within 40 cycles of PCR. <span style='background-color:lightcyan'>Among these specimens, median (interquartile range (IQR)) SARS-CoV-2 Ct was 27.9 (20.4 to 32.9). No secular trend was observed between calendar time from local onset of COVID-19 cases and SARS-CoV-2 Ct values during the study period (Spearman correlation 0.18). **Table 1** summarises the SARS-CoV-2 Ct values, as well as those of the specimen quality control, β-actin, and RT-PCR reaction control, MS2 RNA. **Supplemental Figure 1** depicts the distribution of each target and the relationships between their distrubtions.</span>

@@

We considered the possibility that the observed relationship between SARS-CoV-2 and β-actin Ct values might be confounded by respiratory disease severity, but we found no significant association between independent markers of severe respiratory disease and lower β-actin Ct values. However, several limitations of our analysis must be acknowledged. <span style='background-color:lightcyan'>Measures of disease severity are imprecise,</span> and oxygen saturation data and chest CT radiography reports were only available for a small percentage (32.6% and 8.5%, respectively) of our subjects. Subject demographics and medical comorbidities could not be ascertained for subjects, so unmeasured confounders may contribute to the observed association. <span style='background-color:lightcyan'>The impact of comorbid conditions, immunocompromise, and duration of illness on the observed relationship are important areas for future study.</span>

>>>>

* __Specimen Collection, Processing, and RT-PCR Assay__: Specimens were collected during routine clinical practice using a nylon flocked mini-tip swab collected in VTM or saline [@Garnett2020-dl; @Rogers2020-di; @Richard-Greenblatt2021-ql; @Rodino2020-dk]. Healthcare providers obtained samples using CDC guidelines for NP samples collection. Samples were transported to the laboratory at ambient temperature and stored at 4⁰C if not run immediately. All specimens were analyzed at the Hospital of the University of Pennsylvania's Clinical Microbiology Laboratory using an EUA-approved laboratory-developed SARS-CoV-2 assay on the BD MAX system. The multiplex assay was designed to include a specimen quality control (β-actin)[@Singanayagam2020-mg; @Smieja2010-op], an internal processing control to monitor the RT and PCR steps (MS2 phage RNA)[@Dreier2005-ye], and a SARS-CoV-2 target (Orf1ab). Exk TNA2 extraction reagent kits (Becton Dickinson) for the BD MAX open system reagent suite were used for the lab-developed SARS-CoV-2 assay based on a previously described assay [@LeBlanc2020-wa; @Vermeiren2020-pm]. The BD MAX system was set to run type 1 workflow. PCR conditions consisted of a reverse transcriptase step (600s at 58⁰C, 1 cycle), denaturation step (60s at 98⁰C, 1 cycle) and extension steps (10s at 98⁰C followed by 40s at 63⁰C, 40 cycles). Ct values for all three targets (β-actin, MS2 phage RNA) were recorded. <span style='background-color:lightcyan'>The acceptable Ct range for β-actin in our assay was 24-37. Samples that were negative for SARS-CoV-2 and either negative for β-actin or with a β-actin Ct > 37 were reported as invalid.</span>

@@

* __SARS-CoV-2 Detection and Cycle Threshold Range__: Of 1282 tested specimens, 134 were found to have detectable SARS-CoV-2 within 40 cycles of PCR. <span style='background-color:lightcyan'>Among these specimens, median (interquartile range (IQR)) SARS-CoV-2 Ct was 27.9 (20.4 to 32.9). No secular trend was observed between calendar time from local onset of COVID-19 cases and SARS-CoV-2 Ct values during the study period (Spearman correlation 0.18). **Table 1** summarises the SARS-CoV-2 Ct values, as well as those of the specimen quality control, β-actin, and RT-PCR reaction control, MS2 RNA. **Supplemental Figure 1** depicts the distribution of each target and the relationships between their distributions.</span>

@@

We considered the possibility that the observed relationship between SARS-CoV-2 and β-actin Ct values might be confounded by respiratory disease severity, but we found no significant association between independent markers of severe respiratory disease and lower β-actin Ct values. However, several limitations of our analysis must be acknowledged. <span style='background-color:lightcyan'>Measures of disease severity are imprecise,</span> and oxygen saturation data and chest CT radiography reports were only available for a small percentage (32.6% and 8.5%, respectively) of our subjects. Subject demographics and medical comorbidities could not be ascertained for subjects, so unmeasured confounders may contribute to the observed association. <span style='background-color:lightcyan'>The impact of comorbid conditions, immune compromise, and duration of illness on the observed relationship are important areas for future study.</span>
